---
title: "SQL"
output: html_document
---

##Filtering Columns and Rows

With SQl, you can filter Coulmns and Rows by using SELECT and WHERE. Lets look at an example using the Lahman database. We should first load that database, along with the package sqldf. 

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Suppose now I would like to see the homerun totals for the 1927 Yankees. I could write the following query: 

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```



## ORDER BY

With SQL we can order the rows of our output. For instance, suppose we want to see every instance of a player hitting more than 50 homeruns. But you would like the players with the most homeruns at the top. We could do the following:

```{r}
query<- "SELECT playerID,teamID,yearID,HR From Batting
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```
## AGGREGATION

```{r}
query<-"SELECT playerID,teamID,yearID,SO,AB From Batting
WHERE SO<10 and AB>=400
ORDER BY SO"
sqldf(query)
```

```{r}
query<-"SELECT playerID, HR from Batting
WHERE playerID='ruthba01'"
sqldf(query)
```

```{r}
query<-"SELECT playerID,sum(HR) From Batting
WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```

```{r}
query<-"SELECT playerID,sum(HR) From Batting
GROUP BY playerID
HAVING sum(HR)>600
ORDER BY sum(HR) DESC"
sqldf(query)
```

```{r}
query<-"SELECT playerID, avg(HR) From Batting
GROUP by playerID
HAVING avg(HR)>=30
ORDER BY avg(HR) DESC"
sqldf(query)
```



