---
title: "SQL Website"
output: html_document
---

##Filtering Columns and Rows

With SQl, you can filter Coulmns and Rows by using SELECT and WHERE. Lets look at an example using the Lahman database. We should first load that database, along with the package sqldf. 

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Suppose now I would like to see the homerun totals for the 1927 Yankees. I could write the following query: 

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```



## ORDER BY

With SQL we can order the rows of our output. For instance, suppose we want to see every instance of a player hitting more than 50 homeruns. But you would like the players with the most homeruns at the top. We could do the following:

```{r}
query<- "SELECT playerID,teamID,yearID,HR From Batting
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

```{r}
query<-"SELECT playerID,teamID,yearID,SO,AB From Batting
WHERE SO<10 and AB>=400
ORDER BY SO"
sqldf(query)
```
##Aggregation

When using SQL, there might be a time when the data will need to be condensed for the smallest output. Below are examples of how aggregation is utilized. 
```{r}
query<-"SELECT playerID, HR from Batting
WHERE playerID='ruthba01'"
sqldf(query)
```

```{r}
query<-"SELECT playerID,sum(HR) From Batting
WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```

```{r}
query<-"SELECT playerID,sum(HR) From Batting
GROUP BY playerID
HAVING sum(HR)>600
ORDER BY sum(HR) DESC"
sqldf(query)
```

```{r}
query<-"SELECT playerID, avg(HR) From Batting
GROUP by playerID
HAVING avg(HR)>=30
ORDER BY avg(HR) DESC"
sqldf(query)
```

##JOINS

When usuing SQL, there are times when you want to "join" multiple tables together. Below are a few examples of utilizing JOINS.
```{r}
query<-"SELECT nameFirst,nameLast,name,Batting.yearID,Batting.HR 
FROM (Batting INNER JOIN Master 
ON Batting.playerID=Master.playerID) INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```

```{r}
query<-"SELECT nameFirst,nameLast,sum(HR)
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"
sqldf(query)
```

```{r}
query<-"SELECT nameFirst,nameLast,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"
sqldf(query)
```

